\chapter{Introduction}
\label{chap:introduction}

\section{Background}
\label{section:background}

The fitness industry has seen a significant surge in membership, yet the technological infrastructure 
within gyms remains largely stagnant. Most facilities still rely on basic membership check-ins, 
leaving a massive data gap between the front desk and the gym floor. For members, the lack of real-time 
information regarding equipment availability leads to inefficient workouts and "availability uncertainty." 
For owners, the absence of granular data on how floor space is utilized results in suboptimal equipment 
purchasing and layout decisions. With the advancement of Computer Vision and the ubiquity of CCTV systems, 
there is a significant opportunity to transform passive video feeds into actionable spatial intelligence.

\section{Problem Statement}
\label{section:problem-statement}

Current gym environments operate as "black boxes" once a member passes the front desk. This lack of 
transparency creates two critical points of failure:

1. Dynamic Bottlenecks: Without real-time visibility, members experience "availability uncertainty." 
This leads to routine disruption and peak-hour congestion that management cannot see or redirect, 
directly impacting member retention.

2. Static Asset Management: Owners currently make high-stakes capital decisions—such as purchasing 
five new treadmills or removing a power rack—based on anecdotal evidence rather than objective 
occupancy data. This leads to "dead zones" where expensive square footage is underutilized, 
and "over-taxed zones" where equipment wear-and-tear is accelerated.

3. Infrastructure Scalability: Traditional IoT solutions (individual machine sensors) are cost-prohibitive 
and fragile in high-impact environments. There is currently no scalable way to capture floor-wide 
spatial intelligence using existing infrastructure.

\section{Solution Overview}
\label{section:solution-overview}

Prometheus solves these issues by implementing a non-intrusive AI system that utilizes existing 
gym CCTV cameras. The system uses an Edge-to-Cloud architecture where local processing units 
analyze video streams to identify machine states (Free vs. Occupied). This metadata is then 
synchronized to a cloud database, powering a real-time mobile tracker for members and a spatial 
analytics dashboard for management.

Prometheus bridges the gym’s "data gap" by transforming existing CCTV infrastructure into a real-time 
spatial intelligence network. Using a privacy-first Edge-to-Cloud architecture, the system processes 
video feeds locally to detect equipment occupancy without storing sensitive biometric data.

The solution delivers value through two integrated touchpoints:

1. For Management: A centralized analytics dashboard that maps high-traffic "heat zones" and equipment 
duty cycles, enabling data-driven decisions on floor-plan optimization and capital expenditure.

2. For Members: A real-time mobile interface that eliminates "availability uncertainty" by showing 
live equipment status.

\subsection{Features}
\label{subsection:features}

\begin{enumerate}[leftmargin=90pt]
    \item \textbf{Real-time Occupancy Tracker:} A live mobile interface showing which machines are currently in use or available to help members plan their sets.
    \item \textbf{Spatial Heatmaps:} A web-based visual representation of high-traffic areas and underutilized equipment for facility owners.
    \item \textbf{Progress Logger:} A feature for members to record sets, reps, and weights, integrated with visual progress charts and historical data.
    \item \textbf{Workout Routine Generator:} An intelligent routine creator tailored to user goals, physical profile, and real-time equipment availability.
    \item \textbf{Automated Utilization Reports:} Weekly data summaries for owners regarding peak hours, machine popularity, and equipment ROI.
\end{enumerate}

\section{Target User}
\label{section:target-user}
The Prometheus system is designed for two distinct demographics, with a primary focus on facility management and business intelligence:

\begin{itemize}
    \item \textbf{Gym Owners and Facility Managers (Primary Users):} These are the central stakeholders 
    who utilize the system to drive operational efficiency. They require high-level spatial analytics to 
    identify equipment usage patterns, minimize "dead zones," and optimize facility layouts. By leveraging 
    the web-based dashboard, these users transform raw computer vision data into actionable business insights 
    to improve equipment Return on Investment (ROI) and long-term member retention.
    \item \textbf{Gym-goers (Secondary Users):} Individual patrons who benefit from the downstream data 
    generated by the system. These users range from beginners to advanced athletes who utilize the mobile 
    interface to minimize wait times and follow AI-generated routines. While they are frequent interactors 
    with the platform, their experience is an extension of the primary analytical framework established for 
    the facility.
\end{itemize}

\section{Terminology}
\label{section:terminology}
\begin{itemize}
    \item \textbf{Computer Vision (CV):} A field of Artificial Intelligence that enables computers to derive meaningful information from digital images or videos.
    \item \textbf{YOLO (You Only Look Once):} A state-of-the-art, real-time object detection system used by Prometheus for identifying equipment and patrons.
    \item \textbf{Edge Computing:} Processing data near the source (the camera) to reduce latency and enhance privacy by keeping raw video footage local.
    \item \textbf{Region of Interest (ROI):} Specifically defined coordinates within a video frame that correspond to individual pieces of gym equipment.
    \item \textbf{Occupancy Logic:} The algorithmic determination of a "busy" state based on the spatial intersection of "Person" and "Machine" detection boxes.
\end{itemize}